---
title: "Colony Loss"
author: "Abby Bowen"
format: html
editor: visual
---

```{r}
here::here("data", "colony.csv")
library(gtsummary)
library(readr)
library(dplyr)
colony <- read_csv("data/colony.csv")
View(colony)
```

```{r}
#| label: tbl-months
#| tbl-cap: "Colony Losses and Additions By Month"
tbl_summary(
	colony,
	by = months,
	include = c(colony_lost_pct, colony_added, colony_lost))
```

```{r}
linear_model <- lm(colony_lost_pct ~ months + state + year + colony_added + colony_reno_pct, data = colony)
average_colonies_2020 <- colony %>%
  filter(year == 2020, state %in% c("Georgia", "Minnesota")) %>%
  group_by(state) %>%
  summarise(average_colony_lost = mean(colony_lost, na.rm = TRUE))

stats <- list(n = nrow(colony),
              mean_colony = mean(colony$colony_lost, na.rm = TRUE),
              average_colonies_2020_GA = average_colonies_2020$average_colony_lost[average_colonies_2020$state == "Georgia"],
              average_colonies_2020_MN = average_colonies_2020$average_colony_lost[average_colonies_2020$state == "Minnesota"])
print(stats)
```

```{r}
#| label: tbl-regression
#| tbl-cap: "Multivariable Regression Results"
tbl_regression(
	linear_model,
	exponentiate = FALSE,
	label = list(
		months ~ "Months",
		state ~ "State",
		year ~ "Year",
		colony_added ~ "Colony Added",
		colony_reno_pct ~ "Colony Renovated (%)"
	)
)
```

This data (source: USDA) contains information on honeybee colonies including number of colonies, maximum colonies, number of colonies lost, percent lost, number added, number renovated, and percent renovated. @tbl-months shows seasonal variation in the number of colonies lost, added, and percent lost across the US. Across all states the average number of colonies lost in a given seasonal period is `r stats$mean_colony` (with 4 seaonal periods per year). The avarage number of colony losses in the state of Georgia during 2020 was `r stats$average_colonies_2020_GA`and the number of colonies lost in Minnesota during 2020 was `r stats$average_colonies_2020_MN`

```{r}
#| label: fig-hist
#| fig-cap: "Colony Loss Histogram"
hist(colony[["colony_lost_pct"]],
		 main = paste("Histogram of Colony Losses"),
		 xlab = "Percent of Colonies Lost",
		 ylab = "Frequency",
		 col = "pink",
		 border = "black")
```
